\documentclass[master.tex]{subfiles}
 
\begin{document}


Since the project is rather large (\textgreater 10000 Lines of Code spread over $\sim 100$ Files) a sophisticated build system is required and thus CMake has been choosen.
The components of the simulation are divided into CMake-Targets. There are multiple library targets which encapsulate derivatives, algorithms, helper methods and model equations as well as a few executable targets. One of these executable targets will be the desired simulation.\newline
The project folder structure looks like this:\newline

\begin{forest}
  for tree={
    font=\ttfamily,
    grow'=0,
    child anchor=west,
    parent anchor=south,
    anchor=west,
    calign=first,
    edge path={
      \noexpand\path [draw, \forestoption{edge}]
      (!u.south west) +(7.5pt,0) |- node[fill,inner sep=1.25pt] {} (.child anchor)\forestoption{edge label};
    },
    before typesetting nodes={
      if n=1
        {insert before={[,phantom]}}
        {}
    },
    fit=band,
    before computing xy={l=15pt},
  }
[t3g-cmake
 [extern \textit{Containing External Libraries}] 
 [src
   [algorithms
     [derivatives \textit{finite differences schemes etc.}]
     [poisson \textit{FFTW and SOR Solver}]
     [time \textit{Karniadakis Time Stepper}]
   ]
   [core \textit{Log, Matrix-Class, HDF5 Output}]
   [domain \textit{Shear Shifted Grid Domain}]
   [initialization]
   [models
     [isothermal \textit{Equations implementations}]
     [simulations \textit{Simulation Runner}]
   ]
   [performanceTests \textit{Small programs to evaluate various optimizations}]
   [plotting \textit{API for live plotting}]
   [cuda \textit{Cuda Implementations}]
 ]
 [t3g-config \textit{Default config files}]
 [plotter \textit{Small python program for data visualization}]
]
\end{forest}

Each folder that has C++ code has its own CMakeLists.txt where different build properties and the targets itself are defined (for further information about the CMake code refer to the documentation).\newline
This imposes a clear structure on the code base such that it can be understood by future programmers and easily extended to include further models or algorithms.\newline
With the use of CMake it is also clearly defined what libraries are necessary to build and run the simulation. This is a common problem when building software on different systems for examples clusters since not all libraries are present or located where you would expect them. For example on the \textit{gpu3 cluster} that is used later for evaluation the FFTW3 library is not located at the standard location. A simple solution to this problem is to \textit{tell} CMake the updated path using a simple command line option during the build process:
\begin{blockquote}
  \small
  cmake -DFFTW\_ROOT=/opt/ohpc/pub/libs/gnu7/openmpi3/fftw/3.3.7 (...)\newline
  \newline
  \footnotesize{The git project contains a \textit{Readme} file in which the build process is further described.}
\end{blockquote}




\end{document}